---
output:
  pdf_document: default
  html_document: default
---
```{r, include=FALSE}
library(mosaic)
```


# Exam exercise: Logistic regression analysis of Berkely admission data

You may use the combined lecture notes for this module available at
<https://asta.math.aau.dk> to guide you to the relevant methods and R commands
for this exam.

The following table shows the total number of admitted and rejected
applicants to the six largest departments at University of Berkeley in 1973.

|       | Admitted| Rejected|
|:------|--------:|--------:|
|Male   |     1198|     1493|
|Female |      557|     1278|

Use a $\chi^2$-test to check whether the admission statistics for
Berkeley show any sign of gender discrimination. To enter the table
in R you can do:

```{r}
admit <- matrix(c(1198, 557, 1493, 1278), 2, 2)
rownames(admit) <- c("Male", "Female")
colnames(admit) <- c("Admitted", "Rejected")
admit <- as.table(admit)

# Beregning af forventet procentvis
n <- margin.table(admit)
pctGoals <- round(100 * margin.table(admit, 2)/n, 1)

expected_admit_freq <- (1198+557)/n
expected_reject_freq <- (1493+1278)/n

# Beregning af forventet i tal (rækketotal delt med 100, ganget med forventet procent)
admit_expected_male <- ((1198+1493)/100)*38.8
reject_expected_male <- ((1198+1493)/100)*61.2
admit_expected_fem <- ((557+1278)/100)*38.8
reject_expected_fem <- ((557+1278)/100)*61.2

# Opsætning af forventet tabel for at tjekke overensstemmelse mellem den og funktionens output.
expected_admit <- matrix(c(admit_expected_male, reject_expected_male,
                           admit_expected_fem, reject_expected_fem), 2, 2, byrow = TRUE)
rownames(expected_admit) <- c("Male", "Female")
colnames(expected_admit) <- c("Admitted", "Rejected")
expected_admit <- as.table(expected_admit)

# Chi^2-test
teststat <- chisq.test(admit, correct = FALSE)
round(teststat$expected, 0)
teststat
```

Your analysis should as a minimum
contain **arguments** that support: 

- Statement of hypotheses

$H_0:$ De to grupper, mænd og kvinder, er statistisk uafhængige.

$H_1:$ De to grupper, mænd og kvinder, er statistisk afhængige.

- Calculation of expected frequencies

$\pi_{admit} = 0.388$

$\pi_{reject} = 0.612$

- Calculation of test statistic

$\chi^2 = \sum \frac{(f_0-f_e)^2}{f_e}$

$\chi^2 = 92.205$

- Calculation and interpretation of p-value.

```{r}
p <- 1 - pdist("chisq", 92.205, df = 1, return = "value", plot = FALSE)
p
```

$p = 2.2 \cdot 10^{-16} \approx 0$

p-værdien er praktisk talt lig 0, hvorfor nulhypotesen forkastes, og det må konkluderes, at der er en signifikant statistisk afhængighed mellem de to grupper.



A more detailed data set with the admissions for each department is 
available on the course web page. The variables are:

- `Gender` (male/female)
- `Dept` (department A, B, C, D, E, F)
- `Admit` (frequency of admitted for each combination)
- `Reject` (frequency of rejected for each combination)

Load the data into RStudio:
```{r }
admission <-
    read.table("http://asta.math.aau.dk/dan/static/datasets?file=admission.dat",
               header=TRUE)
admission
```
In order to do logistic regression for this kind of data, the response is the columns `Admit` and `Reject` (which
means that we model the probability of admit) :

```{r }
m0 <- glm(cbind(Admit, Reject) ~ Gender + Dept, family = binomial, data = admission)
```
The glm-object `m0` is a logistic model with main effects of `Gender`
and `Department`.

- Investigate whether there is any effect of these predictors.

```{r}
mainEffects <- glm(cbind(Admit, Reject)~ ., data=admission, family=binomial)
noEffects <- glm(cbind(Admit, Reject) ~ 1, data=admission, family=binomial)

anova(mainEffects, noEffects, test = "Chisq")
```

Med en p-værdi på praktisk talt 0, kan vi kan forkaste $H_0: \beta_1 = \beta_2 = \ldots = \beta_n = 0$



As a hint you might look at section 9.3 in the combined lecture notes.

```{r }
summary(m0)
```

Looking at the summary of `m0`:

- Is there a significant gender difference?

Med en p-værdi på $0.217$ kan nulhypotesen ikke forkastes, hvorfor køn ikke har en signifikant betydning for at blive optaget på universitetet.


- What is the interpretation of the numbers in the `DeptB`-row?

Med en p-værdi på $0.693$ er denne variabel heller ikke signifikant for modellen.


We add the standardized residuals to `admission`:

```{r }
admission$stdRes <- round(rstandard(m0),2)
admission
```
- Looking at the standardized residuals, which department deviates
  heavily from the model?

Department A afviger mest fra de forventede værdier under nulhypotesen.
  
  
- What gender is discrimated in this department?

Det standardiserede residual for mænd i Department A er negativ, hvorfor andelen af mænd, der bliver optaget på denne afdeling er væsentlig lavere end forventet under nulhypotesen. Derfor bliver der diskrimineret mod mænd i denne Department.

  
Next you should fit the model with the interaction `Gender*Dept` and
use `anova` to compare this to `m0`.

```{r}
m1 <- glm(cbind(Admit, Reject) ~ Gender * Dept, family = binomial, data = admission)

anova(m0, m1, test = "Chisq")
```


- Explain what interaction means in the current context.

Det betyder, at der er en interaktion mellem afdeling og køn. Eksempelvis er der generelt flere mænd, der søger ind på STEM-uddannelser, og flere kvinder, der søger ind på sygeplejerske-uddannelsen, hvorfor uddannelsen må have en indflydelse på kønnet af ansøgeren.


- Is there a significant interaction?

Ja.


- In the light of your analysis, explain the reason for your
  answer to the previous question.

p-værdien er lavere end signifikansniveauet, $\alpha = 0.05$. Altså har uddannelsen en indflydelse på kønnet af ansøgeren.


















