---
output: html_document
---

# Exam exercise: Probalilities, CLT and boxplots

It is highly recommended that you answer the exam using Rmarkdown
(you can simply use the exam Rmarkdown file as a starting point).

# Part I: Estimating probabilities

Remember to load the `mosaic` package first:
```{r message=FALSE}
library(mosaic)
library(e1071)
options(digits = 4)
```

## EU climate data

In a recent survey from Eurobarometer you can extract data for
response to the following question:

*Do you consider climate change to be the single most serious problem facing the world as a whole?*

The data are divided according to whether the respondent comes from
Denmark or not.

```{r}
climate <- as.table(matrix(c(309,4918,1010,25830),2))
dimnames(climate) <- list(origin=c("Denmark","Rest of EU"),answer=c("Yes","No"))
climate
```


-   Estimate the probability of answering "Yes" to the question.

```{r}
P_Yes = (309 + 4918) / (309 + 1010 + 4918 + 25830)
P_Yes
```


-   Make a 95% confidence interval for the probability of answering "Yes".

```{r}
#Konf +- 1.96 * sqrt(pihat*(1-pihat)/n)
Konf <- 1.96 * sqrt(0.163*(1-0.163)/(309 + 1010 + 4918 + 25830))
Konf

```


-   Estimate the probability of answering "Yes" given that you come
    from Denmark.
    
```{r}
P_yeS_dk = 309/(309+1010)
P_yeS_dk
```
    

-   What would the true population probabilities satisfy if `origin`
    and `answer` were
    statistically independent? Based on your results do you think they
    are independent? 

```{r}
Hvis B og A er uafhængige
P(A | B) = P(A)
Vi kan se at sandsynligheden for yes ikke er det samme som sandsynlighedenf for yes betinget af at være fra DK
Aka er de afhængige.

```


# Part II: Sampling distributions and the central limit theorem

This is a purely theoretical exercise where we investigate the random
distribution of samples from a known population.

## House prices in Denmark

The Danish real estate agency HOME has a database containing approximately
80,000 house prices for one-family houses under DKK 10 million for the period
2004-2016. The house prices (without all the additional information such as
house size, address etc.) are available as a **R** data file `Home.RData` on the
course webpage. If you download it you can load it using `load("Home.RData")`
assuming you have saved it in the same directory as this Rmarkdown document.
This will add the vector `price` to your work space. Alternatively,
you can add it directly from the course website (this will download it every time
you run the Rmarkdown document, so make sure you have a decent internet connection):

```{r}
load(url("https://asta.math.aau.dk/datasets?file=Home.RData"))
```

Make a histogram of all the house prices using a command like
`gf_histogram(~price, bins = 30)` inserted in a new code chunk
(try to do experiments with the number of bins):

```{r}
his <- gf_histogram(~ price, bins = 50)
his
```


- Explain how a histogram is constructed.
Tæller forekomst af hver pris og placerer i en bin. x aksen er bins og y er antallet af observationen i hver bin.
- Does this histogram look like a normal distribution?
Nej, den ser meget højre skæv ud. Se evt knurhår på boksplottet.

```{r}
box <- boxplot(price, horizontal = TRUE)
```


In this database (our population) the mean price is `r mean(price)` and the
standard deviation is `r sd(price)`.

In many cases access to such databases is restrictive and in the following we
imagine that we are only allowed access to a random sample of 40 prices and the
mean of this sample will be denoted `y_bar`.

Before obtaining this sample we will use the Central Limit Theorem (CLT) to
predict the distribution of `y_bar`:

- What is the expected value of `y_bar`?
```{r}
#Y ~ ?(Mu, sigma) -> Y ~ n(mu, sigma/sqrt(n))

mean(price)

```

- What is the standard deviation of `y_bar` (also called the standard error)?

```{r}
sd(price)/sqrt(40)
```

- What is the approximate distribution of `y_bar`?
Den centrale grænseværdi sætning siger at distrubtionen/fordelingen af middelværdier vil tilnærme sig en normalfordeling ved et passende antal observationer. (25-100)*Skævhed^2 -> Cochrans formel


Now make a random sample of 40 house prices and calculate the sample mean:
```{r}
y <- sample(price, 40)
mean(y)
```

Repeat this command a few times. Is each mean price close to what you expected?

Use `replicate` to repeat the sampling 500 times and save each mean value in the
vector `y_bar`:

```{r}
y_bar <- replicate(500, mean(sample(price, 40)))
mean(y_bar)
sd(y_bar)
```

Calculate the mean and standard deviation of the values in `y_bar`.

- How do they match with what you expected?
Meget godt.

- Make a histogram of the values in `y_bar` and add the density curve for the
approximate distribution you predicted previously using `gf_dist`. 
For example if you predicted a normal distribution with
mean 2 and standard deviation 0.25:

```{r}
gf_histogram(..density.. ~ y_bar, bins = 15) %>%
gf_dist("norm", mean = 1.929, sd = 0.2015, col = "red")
```

- Make a boxplot of `y_bar` and explain how a boxplot is constructed.

```{r}
boxplot(y_bar, horizontal = TRUE)
```


# Part III: Theoretical boxplot for a normal distribution
 
Finally, consider the theoretical boxplot of a general normal distribution with
mean $\mu$ and standard deviation $\sigma$, and find the probability of being an
outlier according to the 1.5$\cdot$IQR criterion:

- First find the $z$-score of the lower/upper quartile. I.e. the value of $z$ such that
  $\mu \pm z\sigma$ is the lower/upper quartile.

- Use this to find the IQR (expressed in terms of $\sigma$).

- Now find the $z$-score of the maximal extent of the whisker. I.e. the value of $z$ such that
  $\mu \pm z\sigma$ is the endpoint of lower/upper whisker.

- Find the probability of being an outlier.
