---
output: html_document
---

# Exam exercise: Reading scores and sample size calculation

It is highly recommended that you answer the exam using Rmarkdown
(you can simply use the exam Rmarkdown file as a starting point).

Remember to load the `mosaic` package first:
```{r message=FALSE}
library(mosaic)
```

## Part I: Directed reading activities

An educator conducted an experiment to test whether new directed reading
activities in the classroom will help elementary school pupils improve
some aspects of their reading ability.

She arranged for a third grade class of 21 students
to follow these activities for an 8-week period. A control classroom of
23 third graders followed the same curriculum without the activities. At
the end of the 8 weeks, all students took a Degree of Reading Power
(DRP) test, which measures the aspects of reading ability that the
treatment is designed to improve.


Read in the data:
```{r}
reading <- read.delim("https://asta.math.aau.dk/datasets?file=reading.txt")
head(reading)
```

Histogram
```{r}
split_reading <- split(reading, reading$Group)
reading_control <- split_reading$Control
reading_treated <- split_reading$Treated
hist(reading_control $ Score)
hist(reading_treated $ Score)
```


Use a boxplot to compare the of measurements of `Score` (the student's DRP score) 
for `Treated`(direct reading activities) and `Control` visually.
```{r}
boxplot(reading $ Score ~ reading $ Group, horizontal = TRUE)
```
Konklusion: 


Use `favstats` to make a numerical summary of the measurements for `Treated` and `Control`.

```{r}
favstats(reading $ Score ~ reading $ Group)
```

-   Write down a point estimate of the mean of the DRP score for students
    following the new *directed reading activities* and explain how this
    is calculated.
Et punktestimat af middelværdien for den "treated" gruppe kan aflæses i favstats til at være 51.48. Det udregnes ved formlen $\sum y_i / n_T$

-   Write down a point estimate of the standard deviation of the DRP score for
    this group and explain how this is calculated.
Sd er i følge favstats 11.01 for "treated" og er udregnet på følgende måde: $S_T = \sqrt{\sum(y_i - \bar y )^2 / n_T - 1} $

-   Write down a 95% confidence interval for the mean of the DRP score for this
    group and explain how this is calculated.
```{r}
mean(reading_treated$Score) + c(-1, 1) * 2.086 * sd(reading_treated$Score)/sqrt(21)
```


Use the command `t.test` to compare the mean of the DRP score of the two groups.

```{r}
t.test(reading $ Score ~ reading $ Group)
```

Go through the details of the output from `t.test`. Your analysis must
include an account of

-   What the relevant null hypothesis and the corresponding alternative
    hypothesis is.
Den relevante nulhypotese er at $H_0 : \mu_T = \mu_C  $ samt den alternative hypotese er at $H_1 : \mu_T \ne \mu_C $

-   Choice and calculation of test statistic.
$ t = \bar y_1 - \bar y_2 / \sqrt{s_1^2 / n_1 + s_2^2 / n_2} $ og ifølge t test er den -2.31. T crit er 2.021 ifølge bogen, fordi df = 40, konfidens interval = 95%. t test er altså mindre end t crit og derfor forkastes vores nulhypotese. Ifølge Qdist er t crit 2.025.

```{r}
qdist("t", p = 1 - 0.05/2, df = 37.885)
```




-   Calculation of $p$-value and its interpretation in connection to a
    conclusion of the analysis.
Ifølge t test er p value lig med 0.026. 
```{r}
p_value <- 2 * pdist("t", df = 37.885, q = -2.3109)
p_value
```

-   Calculation and interpretation of a relevant confidence interval.
$\mathrm{KI} = \bar{y_1} - \bar{y_2} \pm t^{test} \cdot \sqrt{(s_1^2/n_1) + (s_2^2/n^2)}$
Det vil altså sige at konfidens intervallet fortæller $\mu_C - \mu_T = [-18, -2] $, og her kan man altså se at 0 ikke er en del af konfidensinterval. Derfor forkaster vi H0.

## Part II: Determining sample size

_In this part there is no dataset to load into R and analyze. You are only supposed to
use R as a calculator when you apply the relevant formulas (which you find towards
the end of the lecture notes for Module 1)._


A study is being planned to estimate the proportion of the Danish population who smokes regularly. How large a sample size is needed to obtain an estimate which is at most
0.05 away from the true proportion with
a confidence of 0.90? A similar study from 2015 found the estimate of the proportion of smokers to be 22.5%.

Vi skal finde sample size altså n
$ \bar \pi +- Z \cdot \sqrt{\bar \pi  \cdot (1 - \bar \pi) / n} $ Hvor vi ved at 
$  Z \cdot \sqrt{\bar \pi  \cdot (1 - \bar \pi) / n}  = 0.05 $, 0.05 da konfidensinterval er 90% altså 10 % i alt og 5 % i hver "side" af grafen. Hvis vi isolerer n i forrige formel fås 
$ n = Z^2 \cdot \bar \pi \cdot (1- \bar \pi) / 0.05^2 $ 

Løsning 1 : Benyt gammel pi altså 0.225
Løsning 2 : Benyt worstcase pi = 0.5 (tegn grafen, parabel, i Excel)

```{r}
Z <- qdist("norm", p = 1 - 0.1/2, plot = FALSE  )
Z
```

Ved gammel $ \bar \pi $
```{r}
n1 = Z^2 * 0.225 * (1 - 0.225) / 0.05^2
n1
```

Ved worstcase
```{r}
n2 = Z^2 * 0.5 * (1 - 0.5) / 0.05^2
n2
```

